Copulas and their potential for ecology

### 2019-10-15 Paper.Rmd
implemented Lawrence's suggestions

Well-developed approaches do exist, however, and are 
applied widely in other fields, 
that provide a description of the dependence 
between two or more variables that can be mathematically proven 
to be complete [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. 
The approaches are based on 
the concept of the *copula*; copula approaches separate the information of a bivariate 
random variable into two distinct parts: the information in the 
marginal distributions (which says nothing about the dependence between the 
variables), and the remaining information (which is solely about
dependence). We here introduce some simple concepts, based on taking ranks,
that relate to copulas and give a conceptual 
flavor of copula ideas to be introduced 
more formally in the section on Background below. Given 
a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
Fig. \ref{fig_cop_pedag1}A-B), information about the structure of the dependence
between $x$ and $y$ can be separated from information contained 
in the marginal distributions by 
considering the bivariate plot of $u_i$ versus $v_i$, where
$u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
$n+1$; and $v_i$ is defined in the same way but using the $y_i$. The rank
of the smallest element in a set is $1$. The $u_i$ and 
$v_i$ are called *normalized ranks* of the $x_i$ and $y_i$; they relate to 
the empirical cummulative distribution function (cdf) of the $x_i$ and $y_i$,
respectively. See, for instance,
Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
A-B, respectively. Ranking makes the marginal distributions of 
the component datasets uniform, isolating the 
dependence structure. Dependence structure and marginals can then be studied 
separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D
relate to copulas in a way to be described in the section on Background below, 
after a definition of copulas
is provided. Note that Pearson correlation, even though it is the most commonly
used measure of relationships between variables, is modified by monotonic
tranformation of the component variables (Appendix 
\ref{SM-pedagog1_details}, Fig. \ref{SM-fig_transf_pearson_spearman}),
and therefore reflects not only dependence information, but also
information on the marginals [@Genest2007]. 
Spearman and Kendall correlations, being rank-based, 
are not influenced by monotonic transformations of the component variables; 
they provide information solely on the dependence between those variables.

@@

We introduce some standard copula families, as examples of copulas and also
because we will fit these families to data later. The bivariate normal
copula family, which consists of copulas of bivariate normal
distributions, is a one-parameter family. The parameter, which we here call $p$,
corresponds to the correlation
of the related bivariate normal distribution, and controls the degree of dependence
between the variables. A normal copula with $p=0.7$ was already introduced (Fig.
\ref{PedagogFigCopThry}A). The formula for the
copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
cdf of a (univariate) standard normal and $\Phi_{2,p}$ is the joint cdf of a
bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
and $p$ in the off diagonal positions. However, formulas for copulas are often not as 
directly informative about copula properties as pdf diagrams; Fig. \ref{figped_N} has 
example pdfs corresponding to bivariate normal copulas for various values of $p$.
Note that the pdfs are symmetric
around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
associations between the two variables in the left and right tails (see
below for a more precise defintion of tail associations). 
The Clayton copula family,
of which we also already pictured an example (Fig.
\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
have stronger left- than right-tail association. The formula is 
$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
less direct intuition than do example pdfs (Fig. \ref{figped_C}). The survival
Clayton family is the symmetric opposite of the Clayton family, showing
stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
family is a two-parameter family, thereby providing more flexibility: for some parameters
BB1 copulas have stronger left- than right-tail association, and for others the reverse
(Fig. \ref{figped_BB1}). See @nelsen2006_copula and @joe2014_dependence for further
information on these and other copula families.

@@

For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
First, we produced normalized ranks 
$u_i$ and $v_i$ as in the Introduction. 
Second, we tested the independence of the $u_i$ and $v_i$
using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
is Kendall's tau for the data [@Genest2007]. We used the 
implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
in the R programming language. 
We tested for independence because our model selection 
algorithms were ineffective if data could not 
be distinguished from independent data 
since many of the copula families we considered include the independent copula. 
If independence could be rejected ($0.05$ significance level), model 
selection proceeded. Third, we fit 16 bivariate copula families (see below for 
information on the families used) to the 
normalized ranks via maximum likelihood. The approach of fitting
copula families to normalized ranks was studied in depth
and recommended by @Genest1995 and @Shih1995. Their estimator of copula
parameters, which we use, was shown to be consistent, asymptotically normal,
and fully efficient at indepencence [@Genest1995]. @Genest2007
recommend carrying out inferences of dependence structures (which was our
goal here) using normalized ranks.
We used the fitting implementation given in `BiCopEst` in `VineCopula`.
Fourth, we obtained AIC and BIC values and accompanying model weights 
$\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection] for each fitted copula. 
`BiCopEst` also 
provided lower- and 
upper-tail dependence of the best-fitting member of each family. 
$\text{AIC}_{\text{w}}$ values were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

@@

Our model selection procedures do not reveal whether tail dependence parameters are 
*significantly* different from $0$ and from each other. 
Furthermore, we caution that, while our model selection results do convincingly show 
non-normal copula structure, model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula 
(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
available from `BiCopGofTest`).
Our alternative nonparametric
approach, results detailed next, provides informaton about significance of tail
associations and asymmetries of tail associations.

The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
C and N data
were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
of the data,
separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
(Fig. \ref{fig_soilCN_nonparam}). Results showed that tail associations of data are stronger in 
both the lower and upper tails than would be expected under a null hypothesis of a normal copula (Fig. \ref{fig_soilCN_nonparam}B-F).
The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
on surrogates, separately in two 
comparisons (Table \ref{tab_soilCN_asym}). Results showed that
upper-tail association was significantly stronger than lower-tail association. These results
are consistent with the model selection results, but go beyond them by providing information
about statistical significance. Thus our results provide an
affirmative answer to Q1 for the soil C and N data; this answer is 
represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
in the right-most box in the middle row.

@@

We present green spruce aphid abundance (Data section; Table \ref{tab_data_info}) results
as an example multivariate analysis.
Independence was rejected for each pair of the 10 sampling locations. 
Best-fitting (lowest-AIC) copulas were non-normal
for the large majority of location pairs (`r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` of `r tab_multivar_summary$Green_spruce_aphid_count$num.pairs` pairs), 
and the AIC
of the normal copula minus the minimum AIC, averaged across 
location pairs, was `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`.
Whereas `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`
would be only a marginally meaningful AIC difference for a single
location pair, it is more meaningful as an average across many
pairs. Relatedly, and illustrating the concept here, 
the chances of getting `r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` or more 
non-normal location pairs if the chances were equal of getting
a normal versus a non-normal result (and taking into account that
the location pair (A,B) will necessarily 
produce the same result as the pair
(B,A)) is low, 
$`r 1-pbinom(tab_multivar_summary$Green_spruce_aphid_count$num.non_N/2-1,45,0.5)`$. 
These results help answer the first part of Q1: non-normal
copula structure appears to be meaningfully common in
these data, even if not universal.
Goodness of fit tests in every case 
failed to reject the hypothesis that the AIC-best copula 
family was also 
an objectively adequate description of 
the data; i.e., the collection of 
copula families we used was sufficiently broad to characterize 
these data. Model-averaged lower- and upper-tail
dependence statistics had $2.5^{th}$ and $97.5^{th}$ quantiles (across
location pairs)
(`r tab_multivar_summary$Green_spruce_aphid_count$LT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$LT_0.975_CI`) and 
(`r tab_multivar_summary$Green_spruce_aphid_count$UT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$UT_0.975_CI`), respectively,
thereby commonly differing 
from what a normal copula would give (i.e., 0), and helping to
answer the second
part of Q1: these data have greater tail dependence 
(lower and upper) than would be expected from a normal-copula null hypothesis. We note however that our model selection 
procedures again do not reveal whether tail-dependence 
parameters are *significantly* different from $0$, and we refer 
the reader to nonparametric results below for information
on significance. 
Model-averaged lower- minus upper-tail dependence statistics 
were positive for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). The chances
of getting `r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
or more positive values, here, under a null hypothesis of equal chances 
for positive and negative values (and again accounting for the fact that location pairs (A,B) and (B,A) will show the same result) was
again low, 
$`r 1-pbinom(sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)/2-1,45,0.5)`$.
Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
than the spatial synchrony of outbreaks. This answers the third part of Q1.

@@

The first mechanism relates to the ideas in the Introduction about 
Liebig's law of the minimum, and to nonlinear influences of environmental
variables on ecological variables.
If an environmental variables influences an ecological variable 
disproportionately strongly in one of its tails, we explored whether the ecological
variable could then exhibit asymmetric tail dependence across space.
Let $E_i(t)$ be an ecological variable 
measured at location $i$ ($i=1,2$) at time $t$.
Assume the dynamics 
$E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where 
the $\delta_i(t)$ are standard-normally distributed 
and independent across time and locations, $a=0.2$,
$b=0.1, -0.1, 0.5, -0.5$ in different simulations, and 
the $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
$\cov(\epsilon_1,\epsilon_2)=0.8$. Thus basic ecological dynamics follow the
very simple AR(1) formulation, influenced by a "regional" environmental factor which
is correlated across locatons ($\epsilon$) and by an entirely local factor ($\delta$). 
The function $g$ describes how the environmental factor $\epsilon_i(t)$ influences
$E_i(t)$. We used $g$ equal to $g_1$ or $g_2$ in different simulations:
$g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise;
$g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
$g_1$ represents environmental effects that negatively impact populations,
and that occur only below a threshold of $\epsilon=0$; 
and $g_2$ represents effects that positively impact populations and 
that occur only above a threshold of $\epsilon=0$.
The values of $b$ selected provide a modicum of exploration of the potential
for ecological dynamics to influence how complex dependencies between
the $E_i(t)$ may arise: 
negative values of $b$ correspond to overcompensating dynamics 
and positive values to undercompensating dynamics; larger $|b|$
means slower return to equilibrium after a disturbance.

@@

Our second mechanism is an extension of the well known Moran
effect, and was summarized conceptually in the Introduction. 
We consider a linear model, as well as 
two parameterizations of a nonlinear population model which includes density dependence
of population growth. For the linear model, let $E_i(t)$ 
again be an ecological variable, $i=1,2$.
We use AR(1) dynamics, 
$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
The environmental noises $\epsilon_i(t)$ were 
standard-normal random variables that were independent for distinct times, $t$, but 
exhibited different kinds of 
dependence across locations in different simulations (see below).
The variable $E$ is very general, and could represent deviations
of a population density from a carrying capacity, deviations of
total plant community biomass from an average value,
flux of a biogeochemical variable such as methane, or other quantities.
The nonlinear population model was $P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$ for $i=1,2$, using $r=0.5$, $K=100$, $\epsilon_i(t)$
as above, and $\sigma=0.1$ or 
$\sigma=1$ in different simulations. 
When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
$K$, where the model equation can be well approximated by a linear equation,
and the nonlinearities of the model therefore have limited influence on dynamics. When
$\sigma=1$, model dynamics are strongly nonlinear because the stochastic 
component of the model causes populations to stray far from the carrying capacity, $K$. 
We refer to these as
the weak-noise and strong-noise cases, though the importance of the noise here
is that, when it is strong, it brings the nonlinearities of the model into play.

@@

Also consistent with hypothesis, copula structure had a substantial effect on
Taylor's law. 
Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
linearity and homoskedasticity, by non-normal copula structure. For normal copula structure 
(i.e., for simulations that gave normal-copula dependence between populatons
in different locations), $p$-values for the linearty
and homoskedasticity tests were roughly uniformly distributed across replicate 
simulations and Taylor's law appeared visually to be a reasonable approximation
of the $\log(v_g)$ versus $\log(m_g)$ relationship (Fig. \ref{fig_spTL}B,E,F). 
Furthermore, quadratic coefficients
and curvature values were close to $0$ (Fig. \ref{fig_spTL}J,K), and root mean squared
errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
But linearity or homoskedasticity were violated more frequently for non-normal 
copula structure
(Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
non-zero (Fig. \ref{fig_spTL}J,K); and root mean squared
errors from the linear regression were much higher (Fig. \ref{fig_spTL}G). Slopes 
and intercepts of the linear regression were also strongly affected by copula
structure (Fig. \ref{fig_spTL}H,I), though some of the effect here was because
linear regressions do not always adequately represent the $\log(v_g)$ versus $\log(m_g)$
relationship when copula structure was not normal.
Thus our results substantiated the hypothesis, at least for the models we used,
that Taylor's law can be influenced by copula structure and asymmetric
tail associations. This is represented in Fig. \ref{MasterFigure} by the 
arrow labelled "Z".

>>>>

Well-developed approaches do exist, however, and are 
applied widely in other fields, 
that provide a description of the dependence 
between two or more variables that can be mathematically proven 
to be complete [@nelsen2006_copula; @joe2014_dependence; @MaiScherer2017]. 
The approaches are based on 
the concept of the *copula*; copula approaches separate the information of a bivariate 
random variable into two distinct parts: the information in the 
marginal distributions (which says nothing about the dependence between the 
variables), and the remaining information (which is solely about
dependence). We here introduce some simple concepts, based on taking ranks,
that relate to copulas and give a conceptual 
flavor of copula ideas to be introduced 
more formally in the section on Background below. Given 
a sample $(x_i,y_i)$, $i=1,2,\ldots,n$ (e.g., 
Fig. \ref{fig_cop_pedag1}A-B), information about the structure of the dependence
between $x$ and $y$ can be separated from information contained 
in the marginal distributions by 
considering the bivariate plot of $u_i$ versus $v_i$, where
$u_i$ is the rank of $x_i$ in the set $x_j$ ($j=1,\ldots,n$), divided by
$n+1$; and $v_i$ is defined in the same way but using the $y_i$. The rank
of the smallest element in a set is $1$. The $u_i$ and 
$v_i$ are called *normalized ranks* of the $x_i$ and $y_i$; they relate to 
the empirical cumulative distribution function of the $x_i$ and $y_i$,
respectively. (We henceforth abbreviate "cumulative distribution function" as "distribution
function" or "cdf".) See, for instance,
Fig. \ref{fig_cop_pedag1}C-D, which show the normalized ranks of 
A-B, respectively. Ranking makes the marginal distributions of 
the component datasets uniform, isolating the 
dependence structure. Dependence structure and marginals can then be studied 
separately. Normalized rank plots such as Fig. \ref{fig_cop_pedag1}C-D
relate to copulas in a way to be described in the section on Background below, 
after a definition of copulas
is provided. Note that Pearson correlation, even though it is the most commonly
used measure of relationships between variables, is modified by monotonic
tranformation of the component variables (Appendix 
\ref{SM-pedagog1_details}, Fig. \ref{SM-fig_transf_pearson_spearman}),
and therefore reflects not only dependence information, but also
information on the marginals [@Genest2007]. 
Spearman and Kendall correlations, being rank-based, 
are not influenced by monotonic transformations of the component variables; 
they provide information solely on the dependence between those variables.

@@

We introduce some standard copula families, as examples of copulas and also
because we will fit these families to data later. The bivariate normal
copula family, which consists of copulas of bivariate normal
distributions, is a one-parameter family. The parameter, which we here call $p$,
corresponds to the correlation
of the related bivariate normal distribution, and controls the degree of dependence
between the variables. A normal copula with $p=0.7$ was already introduced (Fig.
\ref{PedagogFigCopThry}A). The formula for the
copula is $\Phi_{2,p}(\Phi_{1}^{-1}(u),\Phi_{1}^{-1}(v))$, where $\Phi_1^{-1}$ is the inverse of the
cdf of a (univariate) standard normal and $\Phi_{2,p}$ is the joint cdf of a
bivariate normal distribution with mean $(0,0)$ and covariance matrix having $1$s on the diagonal
and $p$ in the off diagonal positions. However, formulas for copulas are often not as 
directly informative about copula properties as pdf diagrams; Fig. \ref{figped_N} has 
example pdfs corresponding to bivariate normal copulas for various values of $p$.
Note that the pdfs are symmetric
around the diagonal line $v=-u+1$ (Fig. \ref{figped_N}), so normal copulas have symmetric 
associations between the two variables in the left and right tails (see
below for a more precise defintion of tail associations). 
The Clayton copula family,
of which we also already pictured an example (Fig.
\ref{PedagogFigCopThry}B), is another one-parameter family. In contrast to normal copulas, Clayton copulas
have stronger left- than right-tail association. The formula is 
$\left[  \reumax(u^{-p}+v^{-p}-1,0)\right]^{-1/p}$, for parameter $p$; though this again provides 
less direct intuition than do example pdfs (Fig. \ref{figped_C}). Higher values of the parameter,
$p$ produce copulas from the Clayton family with higher Kendall or Spearman correlation. The survival
Clayton family is the symmetric opposite of the Clayton family, showing
stronger right- than left tail association (Fig. \ref{SM-figped_SC}). The BB1 copula 
family is a two-parameter family, thereby providing more flexibility: for some parameters
BB1 copulas have stronger left- than right-tail association, and for others the reverse
(Fig. \ref{figped_BB1}). See @nelsen2006_copula and @joe2014_dependence for further
information on these and other copula families.

@@

For bivariate datasets 
$(x_i,y_i)$ for $i=1,...,n$, model selection involved several steps. 
First, we produced normalized ranks 
$u_i$ and $v_i$ as in the Introduction. 
Second, we tested the independence of the $u_i$ and $v_i$
using the statistic $T=\sqrt{(9N(N-1))/(4N+10)}|\tau|$, where $\tau$
is Kendall's tau for the data. @Genest2007 argue that this statistic is approximately standard normally
distributed. We used the 
implementation of this test given in the `BiCopIndTest` in the `VineCopula` package
in the R programming language. 
We tested for independence because our model selection 
algorithms were ineffective if data could not 
be distinguished from independent data 
since many of the copula families we considered include the independent copula. 
If independence could be rejected ($0.05$ significance level), model 
selection proceeded. Third, we fit 16 bivariate copula families (see below for 
information on the families used) to the 
normalized ranks via maximum likelihood. The approach of fitting
copula families to normalized ranks was studied in depth
and recommended by @Genest1995 and @Shih1995. Their estimator of copula
parameters, which we use, was shown to be consistent, asymptotically normal,
and fully efficient at indepencence [@Genest1995]. @Genest2007
recommend carrying out inferences of dependence structures (which was our
goal here) using normalized ranks.
We used the fitting implementation given in `BiCopEst` in `VineCopula`.
Fourth, we obtained AIC and BIC values and accompanying model weights 
$\text{AIC}_{\text{w}}$ and 
$\text{BIC}_{\text{w}}$ [@burnham2003_modelselection] for each fitted copula. 
`BiCopEst` also 
provided lower- and 
upper-tail dependence of the best-fitting member of each family. 
$\text{AIC}_{\text{w}}$ values were used to get
model-averaged lower- and upper-tail dependence values using standard model averaging 
formulas [@burnham2003_modelselection]; likewise for BIC.

@@

Our model selection procedures do not reveal whether tail dependence parameters are 
*significantly* different from $0$ and from each other. 
Furthermore, we caution that, while our model selection results do convincingly show 
non-normal copula structure, model-averaged tail-dependence statistics may have been biased 
because even the lowest-AIC copula 
(`r BivMS_res_Loecke$InfCritRes$copname[which.max(BivMS_res_Loecke$InfCritRes$AICw)]`) 
was a poor fit ($p=$ `r BivMS_res_Loecke$GofRes_CvM` and `r BivMS_res_Loecke$GofRes_KS` for the 
Cramer-von Mises and Kolmogorov-Smirnov goodness-of-fit tests, respectively, to within the precision 
available from `BiCopGofTest`).
Our alternative nonparametric
approach, results detailed next, provides information about significance of tail
associations and asymmetries of tail associations.

The values of the statistics $\cor_{l_b,u_b}$, $\Ps_{l_b,u_b}$ and $\Dsq_{l_b,u_b}$ for the soil
C and N data
were compared to distributions of their values on 1000 Kendall- or Spearman-preserving normal surrogates
of the data,
separately in two comparisons for each of the ranges $(l_b,u_b)=(0,0.1),\ldots,(0.9,1)$ 
(Fig. \ref{fig_soilCN_nonparam}). Results showed that tail associations of data are stronger in 
both the lower and upper tails than would be expected under a null hypothesis of a normal copula (Fig. \ref{fig_soilCN_nonparam}B-F).
The values of the statistics $\cor_{0,0.1}-\cor_{0.9,1}$, $\Ps_{0,0.1}-\Ps_{0.9,1}$ and
$\Dsq_{0.9,1}-\Dsq_{0,0.1}$ for the soil C and N data were also compared to distributions of their values
on surrogates, separately in two 
comparisons (Table \ref{tab_soilCN_asym}). Results showed that
upper-tail association was significantly stronger than lower-tail association; i.e., C and N values are
more related when high than when low. These results
are consistent with the model selection results, but go beyond them by providing information
about statistical significance. Thus our results provide an
affirmative answer to Q1 for the soil C and N data; this answer is 
represented in Fig. \ref{MasterFigure} as a solid outline around "soil C and N"
in the right-most box in the middle row.

@@

We present green spruce aphid abundance (Data section; Table \ref{tab_data_info}) results
as an example multivariate analysis.
Independence was rejected for each pair of the 10 sampling locations. 
Best-fitting (lowest-AIC) copulas were non-normal
for the large majority of location pairs (`r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` of `r tab_multivar_summary$Green_spruce_aphid_count$num.pairs` pairs), 
and the AIC
of the normal copula minus the minimum AIC, averaged across 
location pairs, was `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`.
Whereas `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC`
would be only a marginally meaningful AIC difference for a single
location pair, it is more meaningful as an average across many
pairs. Relatedly, and illustrating the concept here, 
the chances of getting `r tab_multivar_summary$Green_spruce_aphid_count$num.non_N` or more 
non-normal location pairs if the chances were equal of getting
a normal versus a non-normal result (and taking into account that
the location pair (A,B) will necessarily 
produce the same result as the pair
(B,A)) is low, 
$`r 1-pbinom(tab_multivar_summary$Green_spruce_aphid_count$num.non_N/2-1,45,0.5)`$. 
The average AIC difference `r tab_multivar_summary$Green_spruce_aphid_count$Avg.N_minus_bestfit.AIC` was
much less than the AIC difference between `r round(min(BivMS_res_Loecke$InfCritRes$AIC),2)` and 
`r round(BivMS_res_Loecke$InfCritRes$AIC[1],2)` that was obtained for the soil C and N data
at least in part because the soil C and N data were much more numerous (Table \ref{tab_data_info}).
These results help answer the first part of Q1: non-normal
copula structure appears to be meaningfully common in
these data, even if not universal.
Goodness of fit tests in every case 
failed to reject the hypothesis that the AIC-best copula 
family was also 
an objectively adequate description of 
the data; i.e., the collection of 
copula families we used was sufficiently broad to characterize 
these data. Model-averaged lower- and upper-tail
dependence statistics had $2.5^{th}$ and $97.5^{th}$ quantiles (across
location pairs)
(`r tab_multivar_summary$Green_spruce_aphid_count$LT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$LT_0.975_CI`) and 
(`r tab_multivar_summary$Green_spruce_aphid_count$UT_0.025_CI`, `r tab_multivar_summary$Green_spruce_aphid_count$UT_0.975_CI`), respectively,
thereby commonly differing 
from what a normal copula would give (i.e., 0), and helping to
answer the second
part of Q1: these data have greater tail dependence 
(lower and upper) than would be expected from a normal-copula null hypothesis. We note however that our model selection 
procedures again do not reveal whether tail-dependence 
parameters are *significantly* different from $0$, and we refer 
the reader to nonparametric results below for information
on significance. 
Model-averaged lower- minus upper-tail dependence statistics 
were positive for all but a few location pairs (`r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
out of `r (dim(RES_aphid_count$gfc_p_CvM)[1]^2-dim(RES_aphid_count$gfc_p_CvM)[1])-2*RES_aphid_count$num_indep_loc_pair`). The chances
of getting `r sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)` 
or more positive values, here, under a null hypothesis of equal chances 
for positive and negative values (and again accounting for the fact that location pairs (A,B) and (B,A) will show the same result) was
again low, 
$`r 1-pbinom(sum((RES_aphid_count$LTdep_AICw-RES_aphid_count$UTdep_AICw)>0,na.rm=T)/2-1,45,0.5)`$.
Thus the spatial synchrony of rarity in the green spruce aphid is stronger 
than the spatial synchrony of outbreaks. This answers the third part of Q1.

@@

The first mechanism relates to the ideas in the Introduction about 
Liebig's law of the minimum, and to nonlinear influences of environmental
variables on ecological variables.
If an environmental variables influences an ecological variable 
disproportionately strongly in one of its tails, we explored whether the ecological
variable could then exhibit asymmetric tail dependence across space.
Let $E_i(t)$ be an ecological variable 
measured at location $i$ ($i=1,2$) at time $t$.
Assume the dynamics 
$E_i(t+1)=bE_i(t)+g(\epsilon_i(t))+a \delta_i(t)$, where 
the $\delta_i(t)$ are standard-normally distributed 
and independent across time and locations, $a=0.2$,
$b=0.1, -0.1, 0.5, -0.5$ in different simulations, and 
the $(\epsilon_1(t),\epsilon_2(t))$ were drawn, independently through
time, from a bivariate normal distribution with $\var(\epsilon_i)=1$ and 
$\cov(\epsilon_1,\epsilon_2)=0.8$. Thus basic ecological dynamics follow the
very simple AR(1) formulation, influenced by a "regional" environmental factor which
is correlated across locations ($\epsilon$) and by an entirely local factor ($\delta$). 
The function $g$ describes how the environmental factor $\epsilon_i(t)$ influences
$E_i(t)$. We used $g$ equal to $g_1$ or $g_2$ in different simulations:
$g_1(\epsilon)$ equals $\epsilon$ if $\epsilon<0$ and $0$ otherwise;
$g_2(\epsilon)$ equals $\epsilon$ if $\epsilon>0$ and $0$ otherwise. Thus
$g_1$ represents environmental effects that negatively impact populations,
and that occur only below a threshold of $\epsilon=0$; 
and $g_2$ represents effects that positively impact populations and 
that occur only above a threshold of $\epsilon=0$.
The values of $b$ selected provide a modicum of exploration of the potential
for ecological dynamics to influence how complex dependencies between
the $E_i(t)$ may arise: 
negative values of $b$ correspond to overcompensating dynamics 
and positive values to undercompensating dynamics; larger $|b|$
means slower return to equilibrium after a disturbance.

@@

Our second mechanism is an extension of the well known Moran
effect, and was summarized conceptually in the Introduction. 
We consider a linear model, as well as 
two parameterizations of a nonlinear population model which includes density dependence
of population growth. For the linear model, let $E_i(t)$ 
again be an ecological variable, $i=1,2$.
We use AR(1) dynamics, 
$E_i(t+1)=\beta E_{i}(t)+ \sqrt{1-\beta^2} \epsilon_i(t)$, with $\beta=0.5$. 
The environmental noises $\epsilon_i(t)$ were 
standard-normal random variables that were independent for distinct times, $t$, but 
exhibited different kinds of 
dependence across locations in different simulations (see below).
The variable $E$ is very general, and could represent deviations
of a population density from a carrying capacity, deviations of
total plant community biomass from an average value,
flux of a biogeochemical variable such as methane, or other quantities.
The nonlinear population model was a stochastic, multi-habitat-patch version of the Ricer model, 
$P_i(t+1)=P_i(t)\exp \left(r\left(1-P_i(t)/K  \right)+\sigma\epsilon_i(t)\right)$ for $i=1,2$, using $r=0.5$, 
$K=100$, $\epsilon_i(t)$
as above, and $\sigma=0.1$ or 
$\sigma=1$ in different simulations. 
When $\sigma=0.1$, population dynamics stay close to the carrying capacity, 
$K$, where the model equation can be well approximated by a linear equation,
and the nonlinearities of the model therefore have limited influence on dynamics. When
$\sigma=1$, model dynamics are strongly nonlinear because the stochastic 
component of the model causes populations to stray far from the carrying capacity, $K$. 
We refer to these as
the weak-noise and strong-noise cases, though the importance of the noise here
is that, when it is strong, it brings the nonlinearities of the model into play.

@@

Also consistent with hypothesis, copula structure had a substantial effect on
Taylor's law. 
Taylor's law was strongly influenced, and was often even invalidated in its assumptions of 
linearity and homoskedasticity, by non-normal copula structure. For normal copula structure 
(i.e., for simulations that gave normal-copula dependence between populations
in different locations), $p$-values for the linearty
and homoskedasticity tests were roughly uniformly distributed across replicate 
simulations and Taylor's law appeared visually to be a reasonable approximation
of the $\log(v_g)$ versus $\log(m_g)$ relationship (Fig. \ref{fig_spTL}B,E,F). 
Furthermore, quadratic coefficients
and curvature values were close to $0$ (Fig. \ref{fig_spTL}J,K), and root mean squared
errors from the linear regression were relatively small (Fig. \ref{fig_spTL}G).
But linearity or homoskedasticity were violated more frequently for non-normal 
copula structure
(Fig. \ref{fig_spTL}A,C,D,E,F); quadratic coefficients and curvatures were frequently 
non-zero (Fig. \ref{fig_spTL}J,K); and root mean squared
errors from the linear regression were much higher (Fig. \ref{fig_spTL}G). Slopes 
and intercepts of the linear regression were also strongly affected by copula
structure (Fig. \ref{fig_spTL}H,I), though some of the effect here was because
linear regressions do not always adequately represent the $\log(v_g)$ versus $\log(m_g)$
relationship when copula structure was not normal.
Thus our results substantiated the hypothesis, at least for the models we used,
that Taylor's law can be influenced by copula structure and asymmetric
tail associations. This is represented in Fig. \ref{MasterFigure} by the 
arrow labelled "Z".
